\speculake: Hiding Malware in Speculative Execution

### 2018-08-01 paper/related-work.tex
second pass on weird machines

\speculake is a \textit{weird machine}---a machine which executes arbitrary
computation using unintended executions paths, often the result of bugs or
attacker exploits~\cite{weird_machines,bratus2011exploit,weird_exploits}. In
particular \speculake takes advantage of speculative execution and a separate
trigger program to execute otherwise dead code.

Previous weird machines have demonstrated many traditional computer
vulnerabilities such as buffer overflows, format string exploits and
return-to-libc attacks without function calls~\cite{buffer_overflow,
format_string_exploit, shacham2007geometry}.

Additionally, unexpected functionality such as instruction-less
computing~\cite{bangert2013page} enables the demonstration of a Turing complete
weird machine by taking advantage of memory translation tables, page-faults, and
hardware exception handling to effectively compute arithmetic and logical
operations without the processor executing any instructions.

These types of machines are abundant in modern processors~\cite{d2015exploiting}
as multiple threads share system resources cooperatively and optimizations are
often done across the process isolation boundary.

Whereas previous weird machines have demonstrated the capability to execute
potentially malicious code in an unexpected manner, \speculake showcases the
ability to execute dead code within the speculative world.

>>>>

\speculake shares many properties with \textit{weird machines}---a machine which
take advantage of bugs or unexpected idiosyncracies in existing systems to
perform arbitrary
computation~\cite{weird_machines,bratus2011exploit}. In
particular \speculake showcases the ability to execute dead code.

Previous examples of weird machines include traditional vulnerabilities such as
buffer overflows, format string exploits and return-to-libc attacks without
function calls~\cite{buffer_overflow, format_string_exploit,
shacham2007geometry}.

Unexpected functionality such as instruction-less
computing~\cite{bangert2013page} enables the demonstration of a Turing complete
weird machine by taking advantage of memory translation tables, page-faults, and
hardware exception handling to effectively compute arithmetic and logical
operations without the processor executing any instructions.

Features in modern processors such as multiple threads sharing system resources
and optimizations done across the process isolation boundary leads to
opportunities for ``weird machines''~\cite{d2015exploiting}.

Whereas previous weird machines have demonstrated the capability to execute
potentially malicious code in an unexpected manner, \speculake is a new type of
``weird machine'' that takes advantage of speculative execution and uses a
non-complicit trigger program to execute otherwise dead code.
